---
title: "Sex Ratio and the Price of Agricultural Crops in China"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

In this exercise, we consider the effect of a change in the price of
agricultural goods whose production and cultivation are dominated by
either men or women. 

This exercise is based on: Qian, Nancy. 2008. “[Missing Women and the Price of Tea in China: 
 The Effect of Sex-Specific Earnings on Sex 
 Imbalance.](http://dx.doi.org/10.1162/qjec.2008.123.3.1251)” *Quarterly 
 Journal of Economics* 123(3): 1251–85. 

Our data come from China, where centrally planned production targets during the Maoist
era led to changes in the prices of major staple crops. We focus here
on tea, the production and cultivation of which required a large
female labor force, as well as orchard fruits, for which the labor
force was overwhelmingly male.  We use price increases brought on by
government policy change in 1979 as a proxy for increases in
sex-specific income, and ask the following question: Do changes in
sex-specific income alter the incentives for Chinese families to have
children of one gender over another? The CSV data file,
`chinawomen.csv`, contains the variables shown in the table below, with each observation representing a particular Chinese county in a given year. Note that `post` is
an indicator variable that takes takes 1 in a year following the
policy change and 0 in a year before the policy change.
   
------------------------------------------------------------------------------
 Name                Description
 ------------------- ---------------------------------------------------------
 `birpop`            Birth population in a given year
 
 `biryr`             Year of cohort (birth year)
 
 `cashcrop`          Amount of cash crops planted in county
 
 `orch`              Amount of orchard-type crops planted in county
 
 `teasown`           Amount of tea sown in county 
 
 `sex`               Proportion of males in birth cohort
 
 `post`              Indicator variable for introduction of price reforms
------------------------------------------------------------------------------

## Question 1

We begin by examining sex ratios in the post-reform period (that
  is, the period after 1979) according to whether or not tea crops
  were sown in the region.  Estimate the mean sex ratio in 1985, which
  we define as the proportion of male births, separately for
  tea-producing and non-tea-producing regions.  Compute the 95\%
  confidence interval for each estimate by assuming independence
  across counties within a year (We will maintain this assumption
  throughout this exercise).  Furthermore, compute the
  difference-in-means between the two regions and its 95\% confidence
  interval.  Are sex ratios different across these regions?  What
  assumption is required in order for us to interpret this difference
  as causal?
  
  

## Answer 1  
```{r}
chinawomen<-read.csv("chinawomen.csv")
chinawomen_85<-subset(chinawomen,subset = (chinawomen$biryr=="1985"))
MSR_Tea<-subset(chinawomen_85,subset = (chinawomen_85$teasown!=0 ))
MSR_Non_Tea<-subset(chinawomen_85,subset = (chinawomen_85$teasown==0))

est.conf <- function(x, conf.level){
mean.x <- mean(x, na.rm = TRUE)
se.x <- sqrt(var(x,na.rm = TRUE)/ length(x))
ci <- c(mean.x - qnorm(1 - conf.level / 2) * se.x,
mean.x + qnorm(1 - conf.level / 2) * se.x)
final <- c(mean.x, se.x, ci, length(x))
names(final) <- c("mean", "se", "lower.ci", "upper.ci", "n.obs")
return(final)
}

MSR_Tea_Test<-est.conf(MSR_Tea$sex,0.05)
MSR_Tea_Test

MSR_Non_Tea_Test<-est.conf(MSR_Non_Tea$sex,0.05)
MSR_Non_Tea_Test

est<-MSR_Non_Tea_Test[1]-MSR_Tea_Test[1]

diff_est.conf <- function(x,y, conf.level){
  est <- mean(x, na.rm = TRUE)-mean(y, na.rm = TRUE)
  se.x <- sqrt(var(x,na.rm = TRUE)/ length(x)+var(y,na.rm = TRUE)/ length(y))
  ci <- c(est - qnorm(1 - conf.level / 2) * se.x,
  est + qnorm(1 - conf.level / 2) * se.x)
  final <- c(est, se.x, ci, length(x))
  names(final) <- c("mean", "se", "lower.ci", "upper.ci", "n.obs")
  return(final)
}

diff_est.conf(MSR_Non_Tea$sex,MSR_Tea$sex,0.05)

```


There is no significant difference in the sex ratio between the two regions since,the p_value =0.8728 is greater than 0.05  or the confidence interval contains 0. Therefore the hypothesis that there is no significant change in mean sex ratio between the two regions is not rejected.
The hypothesis is on the assumption that the sex ratio between the two regions are independent.


## Question 2

Repeat the analysis in the previous question for subsequent
  years, i.e., 1980, 1981, 1982, ..., 1990.  Create a graph which
  plots the difference-in-means estimates and their 95\% confidence
  intervals against years.  Give a substantive interpretation of the
  plot.


## Answer 2
```{r}
est<-upper<-lower<-rep(NA,11)
year<-c(1980:1990)
conf.level<-0.05
for(i in 1:11){
chinawomen_i<-subset(chinawomen,subset = (chinawomen$biryr==year[i]))
MSR_Tea<-subset(chinawomen_i,subset = (chinawomen_i$teasown!=0))
MSR_Non_Tea<-subset(chinawomen_i,subset = (chinawomen_i$teasown==0))
est[i] <- mean(MSR_Non_Tea$sex, na.rm = TRUE)-mean(MSR_Tea$sex, na.rm = TRUE)
se.x <- sqrt(var(MSR_Non_Tea$sex,na.rm = TRUE)/ length(MSR_Non_Tea$sex)+
          var(MSR_Tea$sex,na.rm = TRUE)/ length(MSR_Tea$sex))
upper[i]<- est[i] + qnorm(1 - conf.level / 2) * se.x
lower[i]<- est[i] -qnorm(1 - conf.level / 2) * se.x

}

plot(year,est,type="b",col="blue",ylim=c(-0.05,0.05),main="Difference in means After and 95% confidence Interval")
lines(year,upper,type="b",col="red")
lines(year,lower,type="b",col="green")
legend(1984,0.05,legend = c("Upper limit","mean difference","Lower Limit"),lwd = c(1,1,1),col =c("red","blue","green"))
abline(h=0)


```
  
 The confidence interval of the difference in means estimate of sex ratio in 1980 does not include 0 and thus, there is significant change in mean sex ratio between  the two regions in 1980. 
  
  
  
## Question 3

Next, we compare tea-producing and orchard-producing regions
  before the policy enactment.  Specifically, we examine the sex ratio
  and the proportion of Han Chinese in 1978. Estimate the mean
  difference, its standard error, and 95\% confidence intervals for
  each of these measures between the two regions. What do the results
  imply about the interpretation of the results given in Question~1?
  
## Answer 3
```{r}
Tea_78<-subset(chinawomen,subset = c(chinawomen$biryr=="1978" & chinawomen$teasown!=0))
Orchard_78<-subset(chinawomen,subset = c(chinawomen$biryr=="1978" & chinawomen$orch!=0))

diff_est.conf(Tea_78$sex,Orchard_78$sex,0.05)
diff_est.conf(Tea_78$han,Orchard_78$han,0.05)

```

The results show that there is no significant difference in the mean sex ratio between the two regions in 1978 ie( The sex ratio between tea producing and orchard producing regions) and (proportion of han chinese in tea producing and orchard producing regions).  

## Question 4

Repeat the analysis for the sex ratio in the previous question
  for each year before the reform, i.e., from 1962 until 1978.  Create
  a graph which plots the difference-in-means estimates between the
  two regions and their 95\% confidence intervals against years.  Give
  a substantive interpretation of the plot.
  
##Answer 4
```{r}
est<-upper<-lower<-rep(NA,17)
year<-c(1962:1978)
for(i in 1:17){
Tea_78<-subset(chinawomen,subset = c(chinawomen$biryr==year[i] & chinawomen$teasown!=0))
Orchard_78<-subset(chinawomen,subset = c(chinawomen$biryr==year[i] & chinawomen$orch!=0))
est[i] <- mean(Tea_78$sex, na.rm = TRUE)-mean(Orchard_78$sex, na.rm = TRUE)
se.x <- sqrt(var(Tea_78$sex,na.rm = TRUE)/ length(Tea_78$sex)+
          var(Orchard_78$sex,na.rm = TRUE)/ length(Orchard_78$sex))
upper[i]<- est[i] + qnorm(1 - conf.level / 2) * se.x
lower[i]<- est[i] -qnorm(1 - conf.level / 2) * se.x

}
plot(year,est,type="b",col="blue",xlim=c(1962,1978),ylim=c(-0.02,0.05),main="Difference in means 1n Before The Policy and 95% confidence Interval")
lines(year,upper,type="b",col="red")
lines(year,lower,type="b",col="green")
legend(1965,0.05,legend = c("Upper limit","mean difference","Lower Limit"),lwd = c(1,1,1),col =c("red","blue","green"))

```

This show that there was no significant change in mean sex ratio of the two regions before the policy enacted in 1979 except in 1964, where was significant change in the mean sex ratio. 

## Question 5

We will adopt the difference-in-differences design by comparing
  the sex ratio in 1978 (right before the reform) with that in 1980
  (right after the reform).  Focus on a subset of counties that do not
  have missing observations in these two years.  Compute the
  difference-in-differences estimate and its 95\% confidence interval.
  Note that we assume independence across counties but account for
  possible dependence across years within each county.  Then, the
  variance of the difference-in-differences estimate is given by:
  
```{r}

counties_before<-unique(chinawomen$admin[chinawomen$biryr == 1978])
counties_after<-unique(chinawomen$admin[chinawomen$biryr == 1980])
data_before<-subset(chinawomen,subset=(chinawomen$biryr == 1978))
data_after<-subset(chinawomen,subset=(chinawomen$biryr == 1980))

 men_before<-data_before$sex[data_before$sex>0.5]
 women_before<-1-data_before$sex[data_before$sex<0.5]
 men_after<-data_after$sex[data_after$sex>0.5]
 women_after<-1-data_after$sex[data_after$sex<0.5]
 
did_est<-(mean(men_after)- mean(men_before))- (mean(women_after)-mean(women_before))
did_est
   


```
  
  
  
  $$
    (\overline{Y}_{{\text tea}, {\text after}} -  \overline{Y}_{{\text tea},
    {\text before}}) - (\overline{Y}_{{\text orchard}, {\text after}} -  \overline{Y}_{{\text orchard},
    {\text before}}) \\
    (\overline{Y}_{{\text tea}, {\text after}} -  \overline{Y}_{{\text tea},
    {\text before}}) + (\overline{Y}_{{\text orchard}, {\text after}} -  \overline{Y}_{{\text orchard},
    {\text before}}) 
  $$
  
  where dependence across years is given by:
  
  $$
    (\overline{Y}_{{\text tea}, {\text after}} -  \overline{Y}_{{\text tea},
    {\text before}}) \\
    (\overline{Y}_{{\text tea}, {\text after}}) - 2 {\rm
          Cov}(\overline{Y}_{{\text tea}, {\text after}}, \overline{Y}_{{\text tea},
          {\text before}}) + (\overline{Y}_{{\text tea}, {\text before}}) \\
    \frac{1}{n} (Y_{{\text tea}, {\text after}}) - 2 {\rm
          Cov}(Y_{{\text tea}, {\text after}}, Y_{{\text tea},
          {\text before}}) + (Y_{{\text tea}, {\text before}})
  $$
  
  A similar formula can be given for orchard-producing regions.  What
  substantive assumptions does the difference-in-differences design
  require?  Give a substantive interpretation of the results.

