
---
title: "Prediction Based on Betting Markets"
output:
  word_document: default
  pdf_document: default
  
---
  
 
  Earlier in the chapter, we studied the prediction of election outcomes
using polls.  Here, we study the prediction of election outcomes based
on betting markets.  In particular, we analyze data for the 2008 and
2012 US presidential elections from the online betting company, called
Intrade.  At Intrade, people trade contracts such as 'Obama to win
the electoral votes of Florida.'  Each contract's market price
fluctuates based on its sales.  Why might we expect betting markets
like Intrade to accurately predict the outcomes of elections or of
other events?  Some argue that the market can aggregate available
information efficiently.  In this exercise, we will test this 
*efficient market hypothesis* by analyzing the market prices of
contracts for Democratic and Republican nominees' victories in each
state.

The data files for 2008 and 2012 are available in CSV format as
`intrade08.csv` and `intrade12.csv`, respectively. The variables in these datasets are:
  
  --------------------------------------------------------------------------------
  Name                Description
------------------- ----------------------------------------------------------
  `day`               Date of the session 

`statename`         Full name of each state (including District of 
                                             Columbia in 2008) 

`state`             Abbreviation of each state (including District of 
                                                Columbia in 2008)

`PriceD`            Closing price (predicted vote share) of Democratic 
Nominee's market

`PriceR`            Closing price (predicted vote share) of Republican 
Nominee's market 

`VolumeD`           Total session trades of Democratic Party Nominee's 
market

`VolumeR`           Total session trades of Republican Party Nominee's 
market 
--------------------------------------------------------------------------------
  
  Each row represents daily trading
information about the contracts for either the Democratic or
Republican Party nominee's victory in a particular state. 

We will also
use the election outcome data.  These data files are
`pres08.csv` and `pres12.csv` with variables:

--------------------------------------------------------------------------------
Name                 Description
-------------------- ----------------------------------------------------------
`state.name`         Full name of state (only in `pres2008`)

`state`              Two letter state abbreviation

`Obama`              Vote percentage for Obama

`McCain`             Vote percentage for McCain

`EV`                 Number of electoral college votes for this state
--------------------------------------------------------------------------------

We'll also use poll data from 2008 and 2012 in the files `polls08.csv` and `polls12.csv`,  The variables in the polling data are:
  
  --------------------------------------------------------------------------------
  Name                 Description
-------------------- ---------------------------------------------------------
  `state`              Abbreviated name of state in which poll was conducted 

`Obama`              Predicted support for Obama (percentage)

`Romney`             Predicted support for Romney (percentage) 

`Pollster`           Name of organization conducting poll 

`middate`            Middle of the period when poll was conducted 
--------------------------------------------------------------------------------
  
  ## Question 1
  
  We will begin by using the market prices on the day before the
election to predict the 2008 election outcome.  To do this, subset
the data such that it contains the market information for each state
and candidate only on the day before the election.  Note that in
2008 the election day was November 4.  We compare the closing prices
for the two candidates in a given state and classify a candidate
whose contract has a higher price as the predicted winner of that
state.  Which states were misclassified?  How does this compare to
the classification by polls presented earlier in this chapter?
  Repeat the same analysis for the 2012 election, which was held on
November 6.  How well did the prediction market do in 2012 compared
to 2008?  Note that in 2012 some less competitive states have
missing data on the day before the election because there were no
trades on the Republican and Democratic betting markets. Assume
Intrade predictions would have been accurate for these states.

## Answer 1


```{r}
#2008
pres08<-read.csv("pres08.csv",header = T)
intrade08<-read.csv("intrade08.csv",header = T)
polls08<-read.csv("polls08.csv",header = T)

#merge intraday and presidential data
pred<-merge(intrade08,pres08,by="state")

#Find the number of days to the election date
pred$DaysToElection<-as.Date("2008-11-04")-as.Date(pred$day)

#subset the data before the election
data.before<-subset(pred,subset=(pred$DaysToElection==1))

nframe<-data.frame(pres08$state)
#Predicted intrade margin
nframe$win.pred<-data.before$PriceD-data.before$PriceR
# Actual election margin
nframe$win.Elec<-data.before$Obama-data.before$McCain 
#Misclassified states
nframe$pres08.state[sign(nframe$win.pred)!=sign(nframe$win.Elec)]

```


```{r}
#2012
#Follow previous comment
pres12<-read.csv("pres12.csv",header = T)
intrade12<-read.csv("intrade12.csv",header = T)
polls12<-read.csv("polls12.csv",header = T)

pred12<-merge(intrade12,pres12,by="state")

pred12$DaysToElection<-as.Date("2012-11-06")-as.Date(pred12$day)
data.before.12<-na.omit(subset(pred12,subset=(pred12$DaysToElection==1)))#Omits states with missing data.

nframe12<-data.frame(data.before.12$state)
nframe12$win.pred<-data.before.12$PriceD-data.before.12$PriceR
nframe12$win.Elec<-data.before.12$Obama-data.before.12$Romney 

nframe12$data.before.12.state[sign(nframe12$win.pred)!=sign(nframe12$win.Elec)]
```


## Question 2

How do the predictions based on the betting markets change over
time?  Implement the same classification procedure as above on each
of the last 90 days of the 2008 campaign rather than just the day
before the election.  Plot the predicted number of electoral votes
for the Democratic party nominee over this 90-day period.  The
resulting plot should also indicate the actual election result.
Note that in 2008, Obama won 365 electoral votes. Briefly comment on
the plot.

## Answer 2

```{r}

Days<-90:1
sum.obama<-actual<-rep(NA,90)#Set up electoral votes for the predicted and actual for states Obama win.
for(i in 90:1)
{
  data.before<-subset(pred,subset=(pred$DaysToElection==i))#Subset 2008 data on the ith day to the election. 
  data.before$obama.margin<-data.before$PriceD-data.before$PriceR#Calculates the margin
  sum.obama[i]<-sum(data.before$EV[data.before$obama.margin>0])#Sums electoral votes for states Obama is predicted to win using postive margin.
  actual[i]<-sum(pres08$EV[(pres08$Obama-pres08$McCain)>0])#Sums elctoral votes for obama in election
}
sum.obama

#Plots the predicted number of votes for Democratic nominee in the last 90 days to election.
plot(Days,sum.obama,type="l",col="blue",xlab="Days to Election",ylab="Electoral Votes",main="predicted number of votes for Democratic nominee",xlim=c(90,1))
lines(Days,actual,col="red")
legend(0,300,legend = c("Obama","Actual"),col=c("blue","red"),lwd = c(1,1))

```

## Question 3

Repeat the previous exercise but this time use the seven-day
*moving-average* price, instead of the daily price, for each
candidate within a state.  This can be done with a
loop. For a given day, we take the average of the Session Close
prices within the past seven days (including that day).  To answer
this question, we must first compute the seven-day average within
each state. Next, we sum the electoral votes for the states Obama is
predicted to win.  Using the `tapply` function will allow us
to efficiently compute the predicted winner for each state on a
given day.

## Answer 3

```{r}
#2008 intrade
Days<-90:1
w<-7
sum.obama_MA<-actual<-rep(NA,90)
for(i in 90:1)
{
  data.before<-subset(pred,subset = ((pred$DaysToElection<(Days[i]+w)) &      (pred$DaysToElection>=Days[i])))#subset data of 7 day moving avaerage before election 
  obama<-tapply(data.before$PriceD,data.before$state.name,mean)# calculate the moving average for Obama in each state
  mccain<-tapply(data.before$PriceR,data.before$state.name,mean)# Calculate the moving average for McCain in each state
  margin<-obama-mccain #Calculate the margin 
  sum.obama_MA[i]<-sum(pres08$EV[margin>0])#Calculates the number of predicted electoral votes for Obama.
  actual[i]<-sum(pres08$EV[(pres08$Obama-pres08$McCain)>0])#Calculates the actual electoral votes for Obama.
}
sum.obama_MA


plot(Days,sum.obama_MA,type="l",col="blue",xlab="Days to Election",ylab="Electoral Votes",
     main="predicted number of votes for Democratic nominee using MA",xlim=c(90,1))
lines(Days,actual,col="red")
legend(30,300,legend = c("Obama","Actual"),col=c("blue","red"),lwd = c(1,1))


```


## Question 4

Create a similar plot for 2008 state-wide poll predictions using
the data file `polls08.csv`. Notice that polls are not conducted daily
within each state. Therefore, within a given state for each of the
last 90 days of the campaign, we compute the average margin of
victory from the most recent poll(s) conducted. If multiple polls
occurred on the same day, average these polls. Based on the most
recent predictions in each state,sum Obama's total number of
predicted electoral votes. One strategy to answer this question is
to program two loops - an inner loop with 51 iterations for each
state and an outer loop with 90 iterations for each day.

## Answer 4

```{r}
#2008 State Polls

polls<-merge(polls08,pres08,by="state")#Merge polls and election data
polls$DaysToElection<-as.Date("2008-11-04")-as.Date(polls$middate)#Calculates the number of days to election.

state.m<-unique(polls$state)#Create a vector of states.
Days<-90:1
polls.dat<-rep(NA,90)#Creates a vector of predicted number of electoral votes in states Obama won in the last 90 days.
for (j in 90:1)
{
  margin<-rep(NA,51)
  for (i in 1:51) 
  {
  sub.state<-subset(polls,subset = (polls$state==state.m[i]))#subset polls data for the ith state.
  x<-sub.state$DaysToElection #number of days to elction
  sub.polls<-subset(sub.state,subset=(x==min(x[(x-Days[j])>=1])))#Subset data with most recent poll for the ith state
  margin[i]<-mean(sub.polls$Obama.x)-mean(sub.polls$McCain.x) #Calculate the average margin
  }
  ndat<-cbind(pres08,margin)#Election data bind with average margin. 
  polls.dat[j]<-sum(ndat$EV[ndat$margin>0])# Sum the predicted number of electoral votes in the last 90 day
  
}
polls.dat

plot(Days,polls.dat,type="l",col="blue",xlab="Days to Election",ylab="Electoral Votes",ylim=c(200,365),main="predicted number of votes for Democratic nominee using Most Recent Polls",xlim=c(90,1))
lines(Days,actual,col="red")
legend(30,280,legend = c("Obama","Actual"),col=c("blue","red"),lwd = c(1,1))

```

Q.5 What is the relationship between the price margins of the intrade market and the actual margin of victory ? Using only the market data from the day before the election in 2008, regress Obama's actual margin of victory in each state on Obama's price margin from the intrade markets. Similarly, in a seperate analysis, regress Obama's actual margin of victory on the Obama's predicted margin from the latest polls within each state. Interpret the results of the regressions.

```{r}
#2008 actual and intrade regression analysis
#subset the data before the election
data.before<-subset(pred,subset=(pred$DaysToElection==1))

act.margin<-data.before$Obama-data.before$McCain #Actual margin
intr.margin<-data.before$PriceD-data.before$PriceR #Iintrade margin
fit.intrade<-lm(act.margin~intr.margin) #Fitted model of actual margin regressed on intrade margin
summary(fit.intrade)

```

```{r}
#2008 actual and polls regression analysis
#subset the data before the election
data.before<-subset(pred,subset=(pred$DaysToElection==1))

marginpolls<-c() # Creates a vector of margin using polls data before election
for (i in 1:51) 
{
  sub.state<-subset(polls,subset = (polls$state==state.m[i]))#Picks a state
  x<-sub.state$DaysToElection # The number of days to election
  sub.polls<-subset(sub.state,subset=(x==min(x[x>=1]))) # Subset of data with Latest poll in each state
  marginpolls[i]<-mean(sub.polls$Obama.x)-mean(sub.polls$McCain.x) #Find the avaerage margin #obama.x and McCain.x are the polls prediction in the merged data
}
marginpolls<-round(marginpolls,1)
act.margin<-data.before$Obama-data.before$McCain #Actual margin

fit.polls<-lm(act.margin~marginpolls) # Fitted model of actual margin and polls margin
summary(fit.polls)


```

#Q.6
Do the 2008 predictions of polls and intrade accurately predict each state's 2012 elections results? Using the fitted regressions from the previous question,forecast Obama's actual margin of victory for the 2012 election in two ways. First, use the 2012 intrade price margins from the day before the election as the predictor in each state. Recall that the 2012 intradedata do not contain market prices for all states. Ignore states without data. Second, use the 2012 poll predicted margins from the latest polls in each state as the predictor, found in "poll12.csv"

```{r}
#Intrade predicting 2012 margin using 2008 margin regression analysis
pred.2012<-data.before.12#Data of intrade before election omitting states with NA
pred.2012$intr.margin.2012<-pred.2012$PriceD-pred.2012$PriceR#2012 intrade margin
pred.2012$act.margin.2012<-pred.2012$Obama-pred.2012$Romney#2012 election margin

#2012 predicted margin using 2008 regression
pred.2012$pred.12.08<-  1.30265+0.22910 *pred.2012$intr.margin.2012

#Number of states not accurately predicted
data.before.12$state[sign(pred.2012$pred.12.08)!=sign(pred.2012$act.margin.2012)]

```


```{r}
#Polls predicting 2012 margin using 2008 margin regression analysis
polls12<-read.csv("polls12.csv",header = T)
polls.pred.12<-merge(polls12,pres12,by="state")#merge 2012 polls and election data

polls.pred.12$DaysToElection<-as.Date("2012-11-06")-as.Date(polls.pred.12$middate)#Days to election

state.12<-unique(polls.pred.12$state)#selects states which polls were conducted  (there were 47 states in 2012)

marginpolls12<-c() # Creates a vector of margin using latest polls data before election
for (i in 1:length(state.12)) 
{
  sub.state<-subset(polls.pred.12,subset = (polls.pred.12$state==state.12[i]))#Picks a state
  y<-sub.state$DaysToElection # The number of days to election
  sub.polls.12<-subset(sub.state,subset=(y==min(y[y>=1]))) # Subset of data with Latest poll in each state
  marginpolls12[i]<-mean(sub.polls.12$Obama.x)-mean(sub.polls.12$Romney.x) #Find the avaerage margin #obama.x and McCain.x are the polls prediction in the merged data
}

#Creates dataframe of 47 states
data.p<-as.data.frame(state.12)
colnames(data.p)<-"state"

data.p$marginpolls12<-round(marginpolls12,1)#merge 2012 poll margin
polls.12<-merge(polls.pred.12,data.p,by="state")#merge 2012 election data

#2012 predicted polls margin using 2008 regression
polls.12$marginpolls12<- 0.70986 +1.10855 *polls.12$marginpolls12

polls.12$act.margin.2012<-polls.12$Obama.y-polls.12$Romney.y#2012 actual margin

#Number of states not accurately predicted 
polls.12$state[sign(polls.12$marginpolls12)!=sign(polls.12$act.margin.2012)]


```

